---
title: Spring Weather in Minneapolis
author: Andrew Zieffler
date: '2018-04-11'
slug: spring-weather-in-minneapolis
categories: []
tags:
  - R
---

It feels like this spring has been especially terrible weather-wise. We have gotten a lot of snow and it has been cold. To evaluate whether this is the case or whether I have hindsight bias, I pulled some historical weather data for the month of April from [Weather Underground]().

```{r message=FALSE}
library(dplyr)
library(forcats)
library(ggplot2)
library(ggridges)
library(readr)
library(viridis)

# Read in data
april = read_csv("~/Documents/github/Public-Stuff/data/april-weather.csv")

# Filter dates
april = april %>%
  filter(date <= 11)
```

I grabbed data back to 2008 (avialable at https://raw.githubusercontent.com/zief0002/Public-Stuff/master/data/april-weather.csv). To be fair about the comparisons I was making, I filtered the data
so that I only had the first eleven days of April for each year represented in the data. (April is a month in Minnesota that has a lot of variation from the first to last day of the month.) I then counted the number of those days each year that it snowed, as well as, the number of days over 40&#8457;, 50&#8457;, and 60&#8457;.

```{r}
april %>%
  group_by(year) %>%
  summarize(
    days_with_snow = length(grep(pattern = "Snow", x = events)),
    days_over_40 = sum(temp_high > 40),
    days_over_50 = sum(temp_high > 50),
    days_over_60 = sum(temp_high > 60)
  )
```

This suggested that my snow hypothesis (we have had more snow this year) may be erroneous. But, we have had a couple large snows; maybe we had more snow on the days we actually had snow. To evaluate this, I computed the total amount of precipitation for the years it actually snowed during the first eleven days of April.

```{r}
# Get the case numbers for the days that "Snow" is in the event column
snow_days = grep(pattern = "Snow", x = april$events)

# Compute the total amount of precipitation on the days with snow
april %>%
  filter(row_number() %in% snow_days) %>%
  group_by(year) %>%
  summarize(
    total_snow = sum(precip, na.rm = TRUE)
    )
```

While more snow than the last couple years, 2018 doesn't stand out as a banner April for the amount of snow. We will attribute this erroneous hypothesis to snow-blindness.

However, my "April = `r emo::ji("thermometer")`" hypothesis seems spot on. In 2018, so far, we have only had one day where the daily high temperatur was over 40&#8457;. In the 10 years previous, almost all of the first eleven days of April hit over 40&#8457;. This is compounded by the fact that the one day it went over 40&#8457;, the temperature rose to a balmy 41&#8457;!

Lastly, I looked at the distribution of the average daily temperatures since 2008. This was inspired by a post I saw [here](https://cran.r-project.org/web/packages/ggridges/vignettes/gallery.html) and [here](http://austinwehrwein.com/data-visualization/it-brings-me-ggjoy/).

```{r message=FALSE, fig.width=10, fig.height=6}
# Ridge plot of the distribution of average daily temperatures by year
ggplot(april, aes(x = temp_avg, y = fct_rev(factor(year)), fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01, gradient_lwd = 1.) +
  scale_x_continuous(expand = c(0.01, 0)) +
  scale_y_discrete(expand = c(0.01, 0)) +
  scale_fill_viridis(
    name = expression(paste('Temp. (',~degree,'F)',sep='')), 
    option = "C"
  ) +
  labs(
    title = 'April Temperatures in Minneapolis',
    subtitle = 'Mean temperatures for the first 11 days in April',
    x = expression(paste('Average Daily Temperature (',~degree,'F)',sep=''))
  ) +
  theme_ridges(
    font_size = 13, 
    grid = TRUE
  ) + 
  theme(
    axis.title.y = element_blank()
  )
```

This also suggests that we are experiencing a less than tropical April in Minneapolis.

